---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, cache = TRUE,
  comment = "# ",
  fig.path = "tools/README-"
)
```

# ggmap

**ggmap** makes it easy to retrieve raster map tiles from popular online mapping services like [Google Maps](https://developers.google.com/maps/documentation/static-maps/?hl=en), [OpenStreetMap](https://www.openstreetmap.org), [Stamen Maps](http://maps.stamen.com), and plot them using the **ggplot2** framework:

```{r maptypes, message = FALSE}
library(ggmap)

us <- c(left = -125, bottom = 25.75, right = -67, top = 49)
map <- get_stamenmap(us, zoom = 5, maptype = "toner-lite")
ggmap(map)
ggmap(map, extent = "device")
```

Use `qmplot()` in the same way you'd use `qplot()`, but with a map automatically added in the background:

```{r qmplot, message=FALSE}
library(dplyr)

# only violent crimes
violent_crimes <- filter(crime, 
  offense != "auto theft", offense != "theft", offense != "burglary"
)

# rank violent crimes
violent_crimes$offense <- factor(
  violent_crimes$offense,
  levels = c("robbery", "aggravated assault", "rape", "murder")
)

# restrict to downtown
violent_crimes <- filter(violent_crimes,
  -95.39681 <= lon & lon <= -95.34188,
   29.73631 <= lat & lat <=  29.78400
)

qmplot(lon, lat, data = violent_crimes, maptype = "toner-lite", color = I("red"))
qmplot(lon, lat, data = violent_crimes, maptype = "toner-lite", geom = "density2d", color = I("red"))
```

Since **ggmap**'s built on top of **ggplot2**, all your usual **ggplot2** stuff (geoms, polishing, etc.) will work, and there are some unique graphing perks **ggmap** brings to the table, too.

```{r styling, message=FALSE}
robberies <- violent_crimes %>% filter(offense == "robbery")

qmplot(lon, lat, data = violent_crimes, geom = "blank", zoom = 15, maptype = "toner-background", darken = .7, legend = "topleft") +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", alpha = .3, color = NA) +
  scale_fill_gradient2("Robbery\nPropensity", low = "white", mid = "yellow", high = "red", midpoint = 650)
```

Faceting works, too:

```{r faceting, message=FALSE}
qmplot(lon, lat, data = violent_crimes, maptype = "toner-background", color = offense) + 
  facet_wrap(~ offense)
```

For convenience, here's a map of Europe:

```{r europe, message=FALSE}
europe <- c(left = -12, bottom = 35, right = 30, top = 63)
get_stamenmap(europe, zoom = 5) %>% ggmap()
get_stamenmap(europe, zoom = 5, maptype = "toner-lite") %>% ggmap()
```





## Google Maps and Credentials

[Google Maps](http://developers.google.com/maps/terms) can be used just as easily.  However, since Google Maps use a center/zoom specification, their input is a bit different:
```{r google_maps}
get_googlemap("waco texas", zoom = 12) %>% ggmap()
```

Moreover, you can get various different styles of Google Maps with __ggmap__ (just like Stamen Maps):
```{r google_styles}
get_googlemap("waco texas", zoom = 12, maptype = "satellite") %>% ggmap()
get_googlemap("waco texas", zoom = 12, maptype = "roadmap") %>% ggmap()
get_googlemap("waco texas", zoom = 12, maptype = "hybrid") %>% ggmap()
```

Google's geocoding and reverse geocoding API's are available through `geocode()` and `revgeocode()`, respectively:
```{r geocode}
geocode("1301 S University Parks Dr, Waco, TX 76798")
revgeocode(c(lon = -97.1161, lat = 31.55098))
```

There is also a `mutate_geocode()` that works similarly to [__dplyr__](https://github.com/hadley/dplyr)'s `mutate()` function:
```{r mutate_geocode, messages = FALSE}
library(tidyverse)
tb <- data_frame(address = c("1600 Pennsylvania Avenue, Washington DC", "", "waco texas"))
tb %>% mutate_geocode(address)
```


Treks use Google's routing API to give you routes (`route()` and `trek()` give slightly different results; the latter hugs roads):
```{r route_trek}
trek_df <- trek("houson, texas", "waco, texas", structure = "route")
qmap("college station, texas", zoom = 8) +
  geom_path(
    aes(x = lon, y = lat),  colour = "blue",
    size = 1.5, alpha = .5,
    data = trek_df, lineend = "round"
  )
```

(They also provide information on how long it takes to get from point A to point B.)


Map distances, in both length and anticipated time, can be computed with `mapdist()`).  Moreover the function is vectorized:
```{r mapdist}
mapdist(c("houston, texas", "dallas"), "waco, texas")
```



### Google credentialing

If you have a Google API key, you can exceed the standard limits Google places on queries.  By default, when __ggmap__ is loaded it will set the following credentials and limits:
```{r basic_credentials}
ggmap_credentials()
```
Look at the documentation of `?register_google()` to learn more.  If you do have an API key, you set it with:
```{r google_api}
register_google(key = "[your key here]", account_type = "premium", day_limit = 100000)
ggmap_credentials()
```
These will then be used and checked when creating the query URL:
```{r key}
register_google(key = "AbCdEfGhIjKlMnOpQrStUvWxYz")
get_googlemap("waco texas", urlonly = TRUE)
```

For anything that hasn't been implemente (URL-wise), you can inject code into the query usin g `inject`:
```{r otherStuff}
get_googlemap("waco texas", urlonly = TRUE, inject = "otherItem = Stuff")
```






## Installation

* From CRAN: `install.packages("ggmap")`

* From Github: `devtools::install_github("dkahle/ggmap")`

